---
title: "Presenting the {janitor} R package"
author: "LJ"
date: "2022-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.lazy = FALSE,
                      message = FALSE, 
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,
               janitor,
               here,
               readxl,
               kableExtra)

ggplot2::theme_set(theme_minimal())

```

## Introduction

The `janitor` package was created by [Sam Firke](https://samfirke.com/about/)

Its purpose was to make data nomalising as quick and swift as possible, since data cleaning 
is one of the most time consuming activities during data analysis and work.  

This post will go over the `janitor` packages most characteristic functionalities. 

A disclaimer: I had used the package lots of times before, specifically the `clean_names()`
 function.  I will now try to learn of some other ones. 
 
## A messy data example

Data collected or found online is not always normalised, or say, clean.  Data coming 
from excel files must be treated before it is useful for work with programs such as R, or 
Python.  

Some indicators of messyness are:

- strange column or variable names
- empty rows or columns
- consutant number columns
- strange or different date formats

Here there's an example of a messy data set found in the `janitor` github page.  
It's an excel file with data on fictional teachers HR information.

```{r messy data example with kable()}
## This data is visualized using the kable() package
dirty_data <- readxl::read_excel(here::here("data", "dirty_data.xlsx"), skip = 1) # First row containsa an informative cell

dirty_data %>% 
  kbl() %>% 
  kable_classic_2(full_width = FALSE)

```

In the previous view the `kable()` visualization shows us all the data and we can see 
alraedy some of its irregularities:  title rows appear as data rows, the empty columns and rows, etc.

If we view it as a tibble(), we can see also what would make the analysis difficult, and 
the normalization a necessity.  


```{r messy data examples with just tibble()}
dirty_data <- readxl::read_excel(here::here("data", "dirty_data.xlsx"), skip = 1)

dirty_data

```

## Back to Home

[Home](index.html)










































