---
title: "Cleaning titles with clean_names()"
author: "LJ"
date: "2022-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.lazy = FALSE,
                      message = FALSE, 
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,
               janitor,
               here,
               readxl,
               kableExtra)

ggplot2::theme_set(theme_minimal())

```

## Introduction

In the previous session we loaded an example dataset with a lot of noise and 
inconsistencies which we will use in the following:

```{r load messy data example as tibble}
dirty_data <- readxl::read_excel(here::here("data", "dirty_data.xlsx"), skip = 1) # First row containsa an informative cell

dirty_data

```

## Using `clean_names()`

`clean_names()` is used to normalize column names so they follow the same format. 
Unnecessary spaces are removed, all letters become lowercase, and underscores are used
to concatenate and get one single word as a variable.  It also makes the column names readable.

```{r using clean_names()}
d <- dirty_data %>% 
  clean_names()

d

```


```{r}
d %>% 
  glimpse()
```


As we can see, the resulting table uses the first line to create the titles.  Interestingly, 
the real titles are actually found on the second row of values.

Also, duplicate columns get labelled with a number suffix for better differentiation.




## Back to Home

[Home](index.html)
