---
title: "Removing constants and empty row/columns and duplicates"
author: "LJ"
date: "2022-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.lazy = FALSE,
                      message = FALSE, 
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,
               janitor,
               here,
               readxl,
               kableExtra)

ggplot2::theme_set(theme_minimal())

```

## Introduction

In this section we will use the functions to remove constant or empty rows or columns, that 
do not offer any extra informaton, and identify duplicate data in the rows.

```{r add constant columns}
## We will add two constant columns in the dataset d
d <- readxl::read_excel(here::here("data", "dirty_data.xlsx"), skip = 1) %>% 
  clean_names()

d <- d %>% 
  mutate(const_col_01 = 45,
         const_col_02 = "text") %>% 
  relocate(starts_with("const"), .after = last_name)
 

d %>%
   kbl() %>% 
  kable_classic_2(full_width = FALSE)

```

## Remove empty or constant rows/columns

And now let's remove them:

```{r remove the constant columns}
d %>% 
  janitor::remove_constant() %>% 
  janitor::remove_empty() %>% 
  kbl() %>% 
  kable_classic_2(full_width = FALSE)


```

## Identify duplicates


To identify the duplicates in the date we use `get_dupes()`.  This only filters the date.  It does not remove it.  

```{r identify duplicates}

d %>% 
  get_dupes(first_name) %>% 
  kbl() %>%
  kable_classic_2(full_width = F)


```





## Back to Home

[Home](index.html)

























