---
title: "Converting dates"
author: "LJ"
date: "2022-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.lazy = FALSE,
                      message = FALSE, 
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,
               janitor,
               here,
               readxl,
               kableExtra)

ggplot2::theme_set(theme_minimal())

```

## Introduction

Wnen importing data that contains date values, these may come in different formats.
If the import is from excel, these values will not be formatted as dates but as 
numbers.  Excel treats dates as whole numbers starting from 00001 for 1900.

In our dirty_data sample the date column, hire_date has the following values:

```{r show the date column in dirty_data set}
dirty_data <- readxl::read_excel(here::here("data", "dirty_data.xlsx"), skip = 1) %>%   # First row containsa an informative cell
  clean_names()

dirty_data %>% 
  select(hire_date) %>% 
  slice_head(n = 6)

```

## `convert_to_date()`

`janitor` has the `convert_to_date()` function that does the conversion for us.  
What is nice is that it formats the date following the ISO-8601 standard: yyyy-mm-dd


```{r using `convert_to_date()`}
dirty_data %>% 
  select(hire_date) %>% 
  mutate(hire_date = convert_to_date(hire_date))


```


## Back to Home

[Home](index.html)

























