---
title: "Identifying title rows"
author: "LJ"
date: "2022-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.lazy = FALSE,
                      message = FALSE, 
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,
               janitor,
               here,
               readxl,
               kableExtra)

ggplot2::theme_set(theme_minimal())

```

## Introduction

It seems that the janitor package does a lot to keep spreadsheets ready for data
analysis using R.  

It is common in the spreadhseet world to include information in the first 5 rows of a sheet.  
Though it serves the purpose of communicating to the reader and putting them in context, it 
does not do any favors for the data analyst.

Identifying such rows and eliminating from our tables is another cleaning task that can 
be performed by `janitor` using the `row_to_names()` function.

## `row_to_names()` 

```{r create dummy set x}
## Make a dummy dataset
x <- tibble(
  x_01 = c("some general description", "Real title", 1:3),
  x_02 = c("important informatin here", "Another title", 4:6)
)

x

```

In the dummy dataset, x, the real titles are found in row 2.  Passing an argument indicating
the row will eliminate the redundant ones.

```{r using row_to_names}
x %>% 
  janitor::row_to_names(row_number = 2) %>% 
  clean_names()

```


## Back to Home

[Home](index.html)

























