---
title: "Making frequency tables with `tabyl()`"
author: "LJ"
date: "2022-12-04"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE,
                      autodep = TRUE,
                      cache.lazy = FALSE,
                      message = FALSE, 
                      dpi = 180,
                      fig.width = 8, 
                      fig.height = 5)

pacman::p_load(tidyverse,
               janitor,
               here,
               readxl,
               kableExtra)

ggplot2::theme_set(theme_minimal())

```

## Introduction

Whenever a data analysts gets the data to work with, they will need to
do a lot of counting as a first interaction with it, to understand it
and interpret it.

![The Data Science
Lifecycle](here::here(%22docs%22,%22assets%22,%22data-science-explore.png%22))

It is an important of the exploratory stage in data analysis. This is
where we create early graphs (histograms, bar charts, etc.), create
summaries and the like, to make sense of the date or create early
hypothesis from it.

One of the quick ways to see distribution of the data is using frequency
tables.\
Base R has the `table()` function for that; however, it has lots of
drawbacks (not outputting data frames, not easy to format, it doesn't
accept tidy inputs). `janitor` using `tabyl()`, makes this job way more
visual and representative.

## `tabyl()` vignette

For this practice I will use the [`tabyl()`
vignette](https://sfirke.github.io/janitor/articles/tabyls.html#motivation-why-tabyl)
created by Sam Firke, the author of the R package. I find it quite
thorough compared to the blog used in the other sections.

As it starts, the vignette already points to the fact mentioned above
that "data science is mostly counting things."

`tabyl()` is aligned with tidy philosophy, and incorporates functions
that can be used in other formats that are non-tabyl such as tibbles and
data frames. Here, I am referring to the `adorn(_)` family of functions.

`tabyl()` can work with either 1, 2, or 3 input variables (columns) from
a data frame. The next points will go over these ways.

## One-way tabyl

`tabyl()` with one input column will work similarly as Base-R table but
with tidy benefits. Let's use the starwars R package to get a summary of
eye-color from all the human characters.

```{r tabyl with one input column examples: starwars humans}
## Load the starwars package
humans <- dplyr::starwars %>% 
  filter(species == "Human")

humans %>% 
  slice_head(n = 6)


```

```{r tabyl with one input column examples: starwars freq eye color}
t_01 <- humans %>% 
  tabyl(eye_color)

t_01 %>%    # convert as tibble for better visualiztion
  as_tibble()


```

This is similar to doing a count in dplyr, with the added percentage
column in decimals.

If there were NA values, `tabyl()` will show "valid" percentages columns
since it will remove the NAs.

```{r case with NAs}
x <- c("big", "big", "small", "small", "small", NA)
tabyl(x) %>% 
  as_tibble()

```

In the previous example, the input was a vector. As the result is a
tabyl object then it is easy to convert it into workable tidy formats.

The `adorn_()` helper functions may be more suitable for 2-column
inputs; however, they offer more visual info when used in cases like the
ones above.

One these functions is `adorn_pct_formatting()` which adds the % symbol
and roudning to decimal percent columns. `adorn_totals()` adds a total
row if included as argument.

```{r using adorn_pct_formatting and adorn_totals in one-column inputs}
t_01 %>% 
  adorn_totals("row") %>% 
  adorn_pct_formatting() %>% 
  as_tibble()


```

## Two-way tabyl

Frequency tables that include two-column inputs are called crosstab or
contingency tables. Their result is the combination of a dplyr count and
pivot wider.

```{r 2-column input tabyl example}
## Let's create a contingency table of male/femail starwars characters based on eye color
t_02 <- humans %>% 
  tabyl(gender,
        eye_color) %>% 
  as_tibble()

t_02


```

Now let's the percentages by row (contingency tabyls do not output
percentages). Also we will add the corresponding formatting.

```{r adding percents and formatting to 2-column inputs}
t_02 %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns(position = "front") %>%  # Adds the numbers next to the percentages in parenthesis.  Default position is "rear"
  as_tibble()


```

## Three-way tabyls

When adding three input columns, `tabyl()` will produce a list of
tabyls.

In this example we will have the gender show the corresponding
distribution within eye_color, and skin_color in the starwars dataset.

```{r 3-column input example}
t_03 <- humans %>% 
  tabyl(eye_color,
        skin_color,
        gender)

t_03

```

If we want to add `adorn_()` functions to a list of tabyls (or data
frames), the `tabyl()` package will call `purrr_map()` silently to apply
it to each individual element of the list.

```{r}
## Adding adorn_() function to include total rows, percentages, etc.
humans %>% 
  tabyl(eye_color, skin_color, gender, show_missing_levels = FALSE) %>% 
  adorn_totals("row") %>% 
  adorn_percentages("all") %>% 
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns(position = "front") %>% 
  adorn_title()

```

## More on the `adorn_()` functionalities

`adorn_()` will approximate Pivot Table functionalities of Excel on data
frames. Also, they print elegrant results for interactive analysis or
for report sharing.

```{r using adorn_() functionalities for elegant printing}
humans %>% 
  tabyl(gender, eye_color) %>% 
  adorn_totals(c("row", "col")) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(rounding = "half up", digits = 0) %>% 
  adorn_ns(position = "front") %>% 
  adorn_title("combined") %>% 
  knitr::kable()


```

The `adorn_()` functions are...

  a. `adorn_totals()`: Adds totals to row, column or both
  b. `adorn_percentaage()`: Calculates percentages on either rows or columns 
  c. `adorn_pct_formatting()` Format percentage columns, controlling number of digits, 
  and appends the % symbol
  d. `adorn_rounding()`: Round numbers, recommended use of the half_up argument when 
  the decimals land on a .5 (ties situation).  This function should not be used together 
  with `adorn_pct_formatting()`.
  e. `adorn_ns()`: Adds the numbers the percentages represent coming for the underlying 
  counts in the tabyl.  Interesting to see how they can be applied by the user.
  f. `adorn_title()`: Add a title to a tabyl, or a title to a new row on top of the dataframe 
  or a combination of row and column title.
  
  These functions should be applied in this order.  
  
  
## `adorn_()` functions with other types of data frames

`adorn_()` functions work with other data frames, not only the results when using 
`tabyl()`. 

For example, using the mtcars dataset, we can create a column of totals and get 
the percentages from them.

```{r using mtcars with tabyl()}
mtcars %>% 
  rownames_to_column(var = "cars") %>% 
  as_tibble() %>% 
  adorn_totals("col") %>% 
  adorn_percentages("col") %>% 
  adorn_rounding(digits = 2) %>% 
  slice_head(n = 6)

```

```{r example using adorn_() with results from summarise}
percent_above_165_cm <- humans %>%
  group_by(gender) %>%
  summarise(pct_above_165_cm = mean(height > 165, na.rm = TRUE), .groups = "drop")

percent_above_165_cm %>%
  adorn_pct_formatting() 

## As a follow up, it would be good to see how to add Ns (numbers) to the results below.  

```

## Back to Home

[Home](index.html)













































